---
title: "RMarkdown Homework Assignment"
author: "Julia Rogers"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: false
    code_folding: hide
    
---


# Set-up libraries and data

## Load libraries
```{r load-libraries, message = FALSE}
library(tidyverse)
library(pheatmap)
```



## Load data

```{r load-data}
load("data/Rmarkdown_data.Rdata")
```

# Top 20 significant genes

## Get names of top 20 genes
```{r top-genes}
top20_sigOE_genes <- res_tableOE_tb %>% 
  arrange(padj) %>% 	#Arrange rows by padj values
  pull(gene) %>% 		#Extract character vector of ordered genes
  head(n=20)	
```



## normalized counts for top 20 significant genes
```{r normalize-counts}

top20_sigOE_norm <- normalized_counts %>%
  filter(gene %in% top20_sigOE_genes)
```


## Gathering the columns to have normalized counts to a single column
```{r gather-columns}
gathered_top20_sigOE <- top20_sigOE_norm %>%
  gather(colnames(top20_sigOE_norm)[2:9], key = "samplename", value = "normalized_counts")
gathered_top20_sigOE <- inner_join(mov10_meta, gathered_top20_sigOE)
```



## plot using ggplot2
```{r plot-genes}
ggplot(gathered_top20_sigOE) +
  geom_point(aes(x = gene, y = normalized_counts, color = sampletype)) +
  scale_y_log10() +
  xlab("Genes") +
  ylab("log10 Normalized Counts") +
  ggtitle("Top 20 Significant DE Genes") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5))
```

### Expression levels of the top genes in the control, MOV10 knock down, and MOV10 overexpression cells. Most genes are most highly expressed in the overexpressed cells. 

# Create a heatmap of the differentially expressed genes

## Extract normalized expression for significant genes from the OE and control samples (2:4 and 7:9)
```{r extract-expression}
res_tableOE_tb_sig <- res_tableOE_tb %>%
  filter(padj < 0.05)
```
## Return the normalized counts for the significant DE genes
```{r return-counts}
norm_OEsig <- normalized_counts %>% 
  filter(gene %in% res_tableOE_tb_sig$gene) 

meta <- mov10_meta %>%
column_to_rownames("samplename") %>%
data.frame()
```

## Run pheatmap using the metadata data frame for the annotation
```{r make-heatmap}
pheatmap(norm_OEsig[2:9], 
         cluster_rows = T, 
         show_rownames = F,
         annotation = meta, 
         border_color = NA, 
         fontsize = 10, 
         scale = "row", 
         fontsize_row = 10, 
         height = 20)
```

### Heatmap showing clusters of genes, based on their expression in the different cell lines. The top genes are more highly expressed in the overexpressed cells, while the bottom are more highly expressed in the knock down cells. 

## Session Information including package versions
```{r session-information}
  sessionInfo()
```
